---
title: "3 Custom Hints and Tests"
output:
  html_document:
    toc: yes
---
```{r setup, include="FALSE"}
knitr::opts_chunk$set(eval=FALSE, cache=FALSE)
```

# Customizing Hints and Tests

## Automatic Hints

I put a lot of effort into writing sensible automatic tests and hints for RTutor. Hence, in most cases problem set tasks without customized hints or tests should yield a decent experience for users.

If no custom hint is provided, RTutor calls automatically either `hint.for.assign` (if a variable is assigned), `hint.for.call` (function call without assignment), `hint.for.function` (if the user generates a function) or `hint.for.compute` (for multi-step computations in a compute block).

For an example of the automatic hints, consider a chunk were students shall write a dplyr pipe

```s
#< task
dat = tibble(a = sample(1:4,100, replace=TRUE), x=runif(100,0,a))
#>
dat %>%
  group_by(a) %>%
  summarize(mean_x = mean(x)) %>%
  arrange(-mean_x)
```
If the user writes the following wrong code
```s
dat = tibble(a = sample(1:4,100, replace=TRUE), x=runif(100,0,a))
dat %>%
  group_by(a) %>%
  summarize(mean_x = 2*mean(x)) %>%
  arrange(-mean_x)
```
The test fails with a not very informative message:
```
Error: 
You have not yet entered all correct commands in chunk 1.
For a hint, type hint() in the console and press Enter.
```
When typing `hint()` the automatic hint shows the more detailed error analysis:

```
In your following pipe chain, I detect wrong results starting after element 3, the call to 'summarize':

dat %>%
   group_by(a) %>%
   summarize(mean_x = 2 * mean(x)) %>% !! WRONG RESULTS !!
   arrange(-mean_x)
```

There are cases were you may want to customize your hints. Sometimes the automatic hints may reveal to much, sometimes they may reveal to little, or sometimes you just want to really polish your problem set and want hints with nicer text.

There are also (in my experience substantially fewer) cases, in which you want to adapt the automatic correctness checks. We will describe test customization further below.

## Analyse students' submissions to improve hints

In general it is hard to predict where and how hints should be customized. If you use your problem sets in a course in which students submit there solution, take a look at the README.md of the companion package [RTutorSAGI](https://github.com/skranz/RTutorSAGI). It describes how you can analyse log files contained in the submission to detect and improve problematic tasks were many students got stuck. This allows for a nice iterative improvement process for your problem sets.

# Hints

## Custom hints

Consider the following specification in a problem set solution file. The user shall store the mean of a variable `x` in the variable `m` and then show `m`.

```s
m = mean(x)
#< hint
cat("Use Google, e.g. search for 'R compute mean'.")
#>
m
#< hint
cat("Great, you computed m. But read the task exactly, you shall also show it. Just type m in your code.")
#>
```
The `#< hint` blocks describe the code that is run if the user calls `hint()` after the expression directly above was not solved correctly.

We have a different hint for each of the two expressions. Such a custom hint block overwrites the automatic hints, which won't be shown. A motivation to overwrite the first automatic hint is that it makes the solution too simple by already revealing the function name `mean` (although not the argument `y`). Here the custom hint pushes students to learn the extremely useful method to solve problems when coding: search the internet.

The second custom hint is not really neccessary though, since the automatic hint essentially reveals the same info. We can also set custom hints only for some expressions in a chunk, e.g.
```s
m = mean(x)
#< hint
cat("Use Google, e.g. search for 'R compute mean'.")
#>
m
```

## Adaptive Custom Hints

The hints described above always show the same message, independent of the student's code. Automatic hints already adapt to the student's input and from RTutor version 2019.07.23 it is also more simple to make custom hints adaptive.

Here is an example, where the task is to assign the first 100 square numbers to the variable `z`.
```s
z = 1:100 * 1:100
#< hint
if (true(identical(z,1:10*1:10))) {
  cat("Huh, you made a common mistake. Read the task precisely.
  You shall assign to z the first 100 square numbers,
not only all square numbers that are less or equal to 100")
} else if (true(length(z)!=100)) {
  cat("Your variable z must have 100 elements,
  but in your solution z has", length(z),"elements.")  
} else {
  cat("There is a simple one-line formula to compute the first 100 square numbers.
  Just combine what you have learned in exercise 2 f) and in exercise 3 b).
")
}
#>
```

The code in the hint block will be evaluated in an environment in which all variables defined in earlier solved chunks are known. Also (at least if you use the default tests) the student's current code has been already be evaluated in the hint environment. 

This means whether whether `z` exists in the hint environment depends on whether the user has defined it in her solution for the chunk or not. 
The function `true` is a robust version of `isTRUE` that never throws an error. This means even if `z` does not exist and thus the expression cannot be evaluated, we just get a `FALSE`.

## add_to_hint blocks to combine automatic and custom hints

There are many situations were you want to keep the automatic hint but just add some additional hint. Then write your hint code in an `add_to_hint`block.

Here is an example for an adaptive custom hint in an `#< add_to_hint` block from one of my problem sets:

> Task: Using the command `cbind`, generate the matrix X of explanatory variable whose first column consists of 1 and the second column consists of p.

```s
X = cbind(1,p)
#< add_to_hint
if (exists("x") & !exists("X")) {
  cat("It looks like you assigned the value to 'x' (lowercase), but you shall assign the value to 'X' (uppercase).")
}
#>
```
Looking at the logs of students' solution, it became apparent that many mixed up X with x. The automatic hint did not help for this problem, yet in other cases, it was helpful. So I just wanted to add this specific custom hint to the automatic hint.

## Hints at the beginning of chunk

If you place a hint at the beginning of a chunk, the hint text will be added to all the hints of the individual commands.
```s
#< hint
cat("This hint will be added to the automatic hints
of all commands below.")
#>
x = 1:10
y = x^2
```

# Correctness Tests

I tried my best to automatically test whether the student entered a correct solution or not.

A typical reason for adapting the automatic tests or writing custom tests is when you want to allow several correct solutions for a specific task.

## Specifying parameters of automatic tests

Automatic tests either call `check.assign` (if a value is assigned to a variable), `check.call` (a statement that does not assign a variable),  or `check.function` (if a function is generated). These test function have a number of arguments, that allow to customize the tests. 

A `test_arg` blocks allows you to change the arguments of a default test for the preceding statement. Consider the following example:

```
    plot(x=p,y=q,main="A plot", xlab="Prices")
    #< test_arg
      ignore.arg = c("main","xlab")
      allow.extra.arg=TRUE
    #>
```
The `#< test_arg` block customizes the parameters `ignore.arg` and `allow.extra.arg` of the `check.call` function. The parameter `ignore.arg = c("main","xlab")` means that the student does not have to add these two arguments to the plot function or can use different values. The parameter `allow.extra.arg=TRUE` allows the student to specify additional arguments when calling plot, e.g. specifying a `ylab`. So essentially, it will now only be tested whether the x and y arguments of the plot are correct and any customization of the plot will still be considered a correct solution. See the help of `check.call` for a description of arguments.


## Custom tests

If you need more customization, you can use a `test` block. Unlike a `hint` block you cannot just enter arbitrary code that will be evaluated in an approbriate environment. Instead you should call one or several given test functions like `holds.true`, `check.variable`, `check.assign.with.multiple.sol`, `check.expr`,  `check.regression`.

Consider the following an example:
```
    #' b) Save in the variable u a vector of 4 different numbers
    u = c(3,6,7,99)  
    #< test
    check.variable("u",c(3,6,7,99),values=FALSE)
    #>
```

The automatic test `check.assign` would pass if one of the following two conditions is met:

  - `u` has the same value than in the solution. This means `u=c(2,5,6,98)+1` would also pass as correct solution
  - `u` is generated by an equivalent call as in the solution (equivalent means the function name should be the same and the arguments should have the same value). This is useful if the solution is a call that generates a random variable like `x = runif(1)`.
  
Yet in this example, the automatic test is too restrictive. The student shall just generate some arbitrary vector consisting of 4 numbers. The block

```
    #< test
    check.variable("u",c(3,6,7,99),values=FALSE)
    #>
```

replaces the automatic test with a test that just checks whether a variable `u` exists, and has the same length and class (numeric or integer) as an example solution `c(3,6,7,99)`.

## An more complex example of custom tests

Here is a more complex example for customized tests and hints.

> Task: Simulate a vector p of T prices that are correlated (but not perfectly correlated) with the weather w but uncorrelated with the demand shock u.

```{r "1 a) 2"}
p = 1.1*c + 0.5*w
#< test
check.variable("p",1.1*c + 0.5*w,values=FALSE)
test.H0.rejected(cor.test(p,w),failure.message="I don't find a significant correlation between p and w (p-value=={{p_value}})")
holds.true(cor(p,w)<1-1e-14, failure.message="p and w shall not be perfectly correlated!")
test.H0(cor.test(p,u),failure.message="I do find a significant correlation between p and u (p-value=={{p_value}}), but they shall be uncorrelated.")
#>
#< hint
display("To make your prices correlated with w, you have to make w appear in the formula of p. To make p not perfectly correlated with p, there must also be other random factors that influence p, like cost c or some newly drawn vector of random price shocks.")
#>
```
We still provide a sample solution `p = 1.1*c + 0.5*w`, but many other solutions are possible here.

A feasible solution must pass all 4 tests in the test block to be accepted. The function `holds.true` is a quite general tool that can be used in custom tests.

The functions `test.H0.rejected`, and `test.H0` are just examples for special purpose convenience test functions.

Note that if you specify a custom test there will be no automatic hint. We thus added a custom hint.

## Testing a function written by the student

Personally, I find it very valuable if students learn to write own functions. However, it is not easy to nicely test functions and students with little programming experience can get stuck easily.

Assume, you would like students to write an OLS function
```s
ols = function(y,X) {
  Xt = t(X)
  invXtX = solve(Xt %*% X)
  beta.hat = invXtX %*% Xt %*% y
  as.numeric(beta.hat)
}
```
Directly asking them to correctly write such a function will probably fail for many students. 

But you could first create one or several chunks, were students first develop the function body for an example. A solved chunk may look like

```s
# Example arguments
x1 = 1:T 
y = 100 + 2*x1 + rnorm(1)
X = cbind(1,x1)

# Code that we will put inside the function
Xt = t(X)
invXtX = solve(Xt %*% X)
beta.hat = invXtX %*% Xt %*% y
as.numeric(beta.hat)
```
Then provide a function stub like
```s
ols = function(y,X) {
  # Insert your code here
}
```
and let students insert their code here.

Here is an example how you might specify the second chunk in your solution file:
```
    #< fill_in
    ols = function(y,X) {
      
      # enter code to compute beta.hat here ...
      
      return(as.numeric(beta.hat))
    }
    #>
    ols <- function(y,X) {
      beta.hat = solve(t(X) %*% X) %*% t(X) %*% y
      as.numeric(beta.hat)
    }
    #< test_arg
      ols(
        c(100,50,20,60),
        cbind(1,c(20,30,15,20))
      )
    #>
    #< add_to_hint
    display("Just insert inside the function ols the code to compute beta.hat from the previous task")
    #>
```
First, we have a `#< fill_in` block that specifies an unfinished function that will be shown to the student. Afterward, we have an example of a correct function `ols`.
Then the `#< test_arg` block specifies parameters for the automatic test `check.function`. The unnamed parameter 

```
    ols(c(100,50,20,60),cbind(1,c(20,30,15,20)))
```
Specifies a test call. `check.function` will run this test call for both the sample solution and the student's solution. The test will only pass if the both versions of the ols function return the same value.
Finally, the `#< add_to_hint` add some information to the automatic hint.

## Testing a function that generates random variables

If you want to check a function that creates random variables, you can compare the results of the student's function and the official solution using the same random seed.

Here is an example:
```
    a) Write a function `runif.square` with parameters n, min and max
    that generates n random variables that are the square of variables
    that are uniform distributed on the interval from min to max.
    
    ```{r}`r ''`
    runif.square = function(n,min,max) {
      runif(n,min,max)^2
    }
    #< test_arg
      with.random.seed(runif.square(n=20,min=4,max=9), seed=12345)
    #>
    ```
```
Our test call is now embedded in the function `with.random.seed` that calls the function with a fixed random seed. Then the automatically called `check.function` only passes if the students function and official solution return the same value when called with the same random seed.

If a function requires simulation of more than one random number, this testing procedure only works if the student draws the random numbers in the same order than the official solution. This means your task should specify already a lot of structure for the function and tell the student not to draw any additional random variables inside the function.


# Compute blocks for computations with many steps

Sometimes you want to ask students to perform computations that will usually require several intermediate steps.
Two somewhat opposite ways of implementing such multistep computations in a problem set would be the following:

  i) Only check the final result and let the student figure out the intermediate steps herself.
  ii) Include all intermediate steps as part of the problem set

A `#< compute ... #>` block allows an intermediate approach. Here is an example from a problem set of mine that asks a student to compute a matrix of choice probabilities from a conditional logit model.

```s
    #< compute P
    
    ## Take a look at the formula for the choice probabilities P.nj of the logit discrete choice model in the slides.
    
    ## Let exp.V be a matrix that contains the numerators of P.nj (use scale as sigma in your formula)
    exp.V = exp(V/scale)
    
    ## Let sum.exp.V be a vector that contains for each person n the denominator in P.nj. You can use the function 'rowSums'. 
    sum.exp.V = rowSums(exp.V)
    
    ## Compute P as ratio of the numerator and the denominator
    P = exp.V / sum.exp.V
    #>
```

The solution will pass as correct if the final values of `P` are correct. The student is not obliged to perform the particular intermediate computations like `exp.V`. Yet, if the student has not yet correctly computed `P` and types `hint()`, the hint function tries to steer the student step by step through the sample solution described in the block. The comments starting with `##` will be transformed into text that will be shown in the hint.
