---
title: "Deploy a problem set on shinyapps.io"
author: "Sebastian Kranz"
date: "15.03.2015"
output:
  html_document:
    highlight: textmate
    theme: readable
    toc: yes
---

This vignette explains how you can host a problem set directly in the web using shinyapps.io. The steps are also useful if you want to host it on another web server that runs shiny server.

You can create an app based on an .rps file or based on a problem set package hosted that is hosted on Github (see the corresponding vignette for how to create such a package)


### 1. Create an app based on an skeleton with `rtutor.app.skel`

The RTutor package has a function `rtutor.app.skel` that creates a skeleton of your app directory and a file `deployapp.r` which contains commands to deploy your app. 

Here is an example fore creating an app based on an .rps file:




```{r eval=FALSE}
  # Create app based on .rps
 
  ps.name = "Example" # problem set name
  # Folder that contains your .rps file
  rps.dir = "D:/libraries/RTutor/examples"
  # Folder in which app shall be created
  app.dir = "D:/libraries/RTutor/examples/ExampleApp"
  rtutor.app.skel(ps.name=ps.name, app.name="RTutorExample",
                  app.dir=app.dir, rps.dir = rps.dir,
                  rps.app = TRUE, overwrite=TRUE)
 
```

Creating an app based on an problem set package works similar. Here is an example:

```{r eval=FALSE}
  # Create app based on a problem set package
  ps.name = "Example"
  app.dir = "D:/libraries/RTutor/examples/ExampleApp"
  rtutor.app.skel(ps.name=ps.name, app.name="RTutorExample",app.dir=app.dir, 
                  pkg.name = "RTutorExample", rps.app = FALSE,
                  github.user = "skranz", overwrite=TRUE)
```

In both cases you must just adapt the examples above.

### 2. Open `deployapp.r` and follow the described steps

Now in your app directory, e.g. "D:/libraries/RTutor/examples/ExampleApp" a file `deployapp.r` is created. Follow the steps described in this file to deploy your app on shinyapps.io. The file also contains the links that describe how to create a user account on shinyapps.io.

