Guide to fixing errors when solving a problem set
========================================================

Here is some advice how to proceed if you get an error when solving your interactive problem set that you don't know how to solve.

## A Seven steps to solve an error in a problem set:

Please follow the seven steps in this order if you encounter an error that you cannot solve.

### 1. Check you solution with Ctrl+S & Ctrl+Alt+R

The whole idea of RTutor is to provide some help in automatically checking your solution. Save your .Rmd file by pressing Ctrl+S and then evaluate your code by pressing Ctrl+Alt+R (running all chunks.) This has the effect that the function `check.problem.set` at the beginning of your code is run, which tries to automatically check your solution.

### 2. Check out for a hint

Sometimes you have the option to type hint() to get a hint of how to solve the exercise.

### 3. Make sure you have installed the newest version of RTutor (from github) and have downloaded the newest version of the .rps file.

Sometimes an error is due to a bug in the problem set or in RTutor which may have already been fixed in the newest versions.

### 4. Take a look at the list of common errors below in this document.

Some errors will not be detected by RTutor or the error message is not really useful. Below is a list that should help you to identify and resolve some of those errors.

### 5. If your course has a discussion forum take check if somebody else had a similar error and whether there is an answer that solves your problem.

It is generally a good idea for course instructors to create such a discussion forum, e.g. on the Moodle site of the course.

### 6. If your course has a discussion forum use it to post a question. Show the R output for the error and maybe a one or two lines of your code where the error occurs. (But *don't* upload your whole problem set.)

On the discussion forum you should get answers either by other students or by the teaching assisstant or lecturer of the course. The nice thing is that other students who encounter similar problems can then also use these answers.

### 7. Only if points 1 to 6 don't help, send an email to your teaching assisstant (and possibily to your lecturer) with your .Rmd file and a description of your error.

## B Frequently encountered errors

### 1. You destroyed the header or footer line of a code chunk

Do you get an error message looking similar to the following:

    > source('~/.active-rstudio-document', echo=TRUE)
    Error in source("~/.active-rstudio-document", echo = TRUE) : 
      ~/.active-rstudio-document:36:2: unexpected ')'
    35: 
    36: b)

It could then be that you have destroyed the header or footer line of a code chunk. Your problem sets have the format of R-Markdown files (.Rmd) which are simple text files that allow to mix formatted text with R code. Here is an example: 

    a) Compute in R 5*5
    ```{r }
    5*5
    ```
    b) Now store the value 10 in the variable x ...


A code chunk starts with a line like
    ```{r }      
and ends with a line
    ```
Never the change those start and end lines!

Such things can happen by accident, however, e.g. you may accidently remove the footer line of the code chunk.

    a) Compute in R 5*5
    ```{r }
    5*5
    
    b) Now store the value 10 in the variable x ...

You will then get a cryptic error message as above. The best is to search for the code shown in the error message (here `b)`) and check whether you messed up a code chunk above.
Note that RStudio shows code chunks with a light-grey background color while text has white background color. This should help to identify such problems.

spin("Guide for fixing errors when solving a problem set.Rmd", knit=FALSE, format="md")
