# Problemset Example

Example of an RTutor interactive Problemset
Author:  Sebastian Kranz
Date:    12.06.2014

Adapt the working directory below and press Ctrl-Alt-R (run all chunks). This creates the problem set files and the sample solution from this solution file.

```{r "Create problem set files"}
library(RTutor)
setwd("C:/folder_of_this_solution_file")
ps.name = "Example"; sol.file = paste0(ps.name,"_sol.Rmd")
libs = NULL # character vector of all packages you load in the problem set
create.ps(sol.file=sol.file, ps.name=ps.name, user.name=NULL,libs=libs, name.sol.chunks=TRUE)
```
 

## Exercise 1 -- Summary statistics #####################################

a) We often want to compute some summary statistic of a vector.
For example:
```{r}
#< task
x = 10:20
# Computing the sum of x
sum(x)
#> task
```
Now compute the mean of x.
```{r}
mean(x)
#< hint
display("There already exist R functions for many things. To find them use Google, e.g. search for 'R compute mean'.")
#>
#<
give.award("mean means mean","Well, in some occasions one can just guess the name of an R function. The function to compute the mean of a vector, or matrix is called 'mean'. Usually, it is much quicker to goggle than to guess function names, however.")
#>
```


## Exercise 2 -- Computing with vectors #####################################

```{r}
#< settings
import.var = list("1"="x")
#>
```
a) Let y be a vector that contains the squared elements of x. Show y
```{r}
y = x^2
```

