# Problemset r_intro

An interactive Problemset for RTutor that introduces to computing with vectors in R

Author:  Sebastian Kranz
         (with help from Beat Classen)
Date:    16.04.2014
License: GPL >= 2.0

## Exercise 1 -- R as calculator

a) Compute 123 + 257 by typing it below as R code. To run the code in your R console, go to the line with your text cursor and press Ctrl-Enter.

```{r "1 a)"}
123+257
#< hint
cat("'Just type 123+257 where the R code shall be entered in part a).")
#>
```

b) The command
```{r "1 b)"}
#< task
x = 4
#> task
```

assigns the value 4 to the **variable** x.

Assign the value 8 to the variable y

```{r "1 b) 2"}
y = 8
#< hint
cat("Come on, you see above that x = 4 assigns the value 4 to the variable x. Do you have no idea how you could assign the value 8 to the variable y?")
#>
```

c) We can now calculate with these variables. Compute the product of x and y

```{r "1 c)"}
x*y
#< hint
cat("a+b would show the sum of a variables a and b.")
#>

#<
give.award("CalculatoR")
#>
```

## Exercise 2 -- Calculating with vectors

a) A nice feature of R is that you can also directly calculate with vectors.
The following command creates a vector of 3 numbers

```{r "2 a)"}
#< task
c(4,2,73)  
#> task
```

b) Save in the variable u a vector of 4 different numbers

```{r "2 b)"}
u = c(3,6,7,99)  
#< test
check.var("u",c(3,6,7,99),exists=TRUE, length=TRUE, class=TRUE)
#>
```

c) Enter the command 3*u

```{r "2 c)"}
3*u
```

The command simply multiples each element of u by 3.

d) Add 100 to each element of u

```{r "2 d)"}
100+u
```

e) Generate another vector v of 4 different numbers and compute the sum of u and v

```{r "2 e)"}
v = 1:4
#< test
check.var("v",1:4,exists=TRUE, length=TRUE, class=TRUE)
#>
u+v
```

f) Find out what the command u*v computes. Is it the scalar product of u and v, the outer product or the element-wise product?

```{r "2 f)"}
#< task
# Write down as answer "scalar product", "outer product", or "element-wise product"
#> task
"element-wise product"


#<
give.award("Vector-CalculatoR")
#>
```

## Exercise 3 -- Combining vectors

a) Consider the following commands that generate vectors

```{r "3 a)"}
#< task
x = 2:4
x
c(x, 10:13)
seq(5,7,by=0.5)
#> task
```

By looking at the resulting vectors can you guess what the commands a:b, c and seq do? You can also use the R-help by pressing F1 when your text-input cursor is on the command name. 

b) Using : create a vector that contains all integer numbers between 10 und 20, i.e. 10, 11, 12, ..., 19, 20.

```{r "3 b)"}
10:20
```

c) Let y be a vector that contains all numbers between 0 and 1 with step size 0.01, i.e. 0, 0.01, 0.02, ..., 0.99, 1. Then show y.

```{r "3 c)"}
y = seq(0,1,by=0.01) 
y
#< hint
cat('
Search in Google for "R generate a sequence of numbers" and check the shown mailing list answers or R help files for a command that can generate such sequences.
Otherwise use the ?seq command to find out, what parameters you can set and what they do.
')
#>
```

d) Let 'b' be a variable that contains all integer numbers from 0 to 10 and from 50 to 60 but not the numbers inbetween. Show b.

```{r "3 d)"}
b = c(0:10, 50:60)  
b
```

e) Let z be a variable that shows the first 100 square numbers, i.e. 1,2,4,9,... Then show z

```{r "3 e)"}
z = 1:100 * 1:100
#< hint
cat("There is a simple one-line formula to compute the 100 first square numbers. Just combine what you have learned in exercise 2 f) and in exercise 3 b).")
#>
z
#<
give.award("V[e]ctory!")
#>
```

## Exercise 4 -- Summary functions

a) We often want to compute some summary statistics of a vector. For example:

```{r "4 a)"}
#< task
x = 10:20
x
# Compute the sum of x
sum(x)
#> task
```

b) Compute the mean of x.

```{r "4 b)"}
mean(x)
#< hint
display("Search Google for something like 'R compute mean' to find the right function name.")
#>
```

c) Compute the standard deviation of x.

```{r "4 c)"}
sd(x)
#< hint
display("Search Google for something like 'R compute standard deviation' to find the right function name.")
#>
```

d) Find the maximum value of all elements in x.

```{r "4 d)"}
max(x)
#< hint
display("Search Google for something like 'R find maximum' to find the right function name.")
#>
```

e) Find the minimum value of all elements in x.

```{r "4 e)"}
min(x)
#< hint
display("Search Google with 'R ...' to find the right function name.")
#>
```

f) Find the index of the element with the maximum value of all elements in x.
Note: There are several ways to find the index of the maximum element. Here you shall use a single function call.

```{r "4 f)"}
which.max(x)
#< hint
display("Search Google with 'R ...' to find the right function name. The function you shall use, has the structure 'which.???'")
#>

#<
give.award("Summarizer")
#>
```

## Exercise 5 -- Indexing

a) The following commands some numbers in a vector x

```{r "5 a)"}
#< task
x = ((1:19)-10)^2
x
```

We can simply show a plot of x by typing

```{r "5 a) 2"}
plot(x)
```

b) If we want to access the 3rd element of x, we  type

```{r "5 b)"}
x[3]
```

We can get a subvector of the elements 5,6,7 by typing

```{r "5 b) 2"}
x[5:7]
#> task
```

c) Show the 9nth element of x

```{r "5 c)"}
x[9]
```

d) Save in the variable z the elements 15,16,17,18 of x and show z

```{r "5 d)"}
z = x[15:18]
z
```

e) Store in y the elements 4,2,9 (in this order) of x and show y

```{r "5 e)"}
y = x[c(4,2,9)]
#< hint
cat('
You can use any vector as index, e.g. x[c(3,2,1)] would yield the elements 3,2,1 of x.')
#>
y
```

## Exercise 6 -- Logical operations

a) R has logical values TRUE or FALSE

```{r "6 a)"}
#< task notest
# Comparisons yield TRUE or FALSE
5 > 3
1 == 2 # == checks equality

x = 10 # = assigns a value
x < 3
```

The operator & is a logical AND. It returns TRUE if BOTH conditions are TRUE

```{r "6 a) 2"}
5 > 3 & x==2 
1+2 == 3 & 100 / 10 == x
```

The operator | is a logical OR. It returns TRUE if at least ONE of the two conditions is TRUE

```{r "6 a) 3"}
5 > 3 | x==2 
1+2 == 3 | 100 / 10 == x
1+1 == 3 | 2+2 == 5
```

We can apply comparisons and logical operators also on vectors of the same length, which are then performed elementwise. We get a logical vector that contains TRUE and FALSE for each element of the vector

```{r "6 a) 4"}
y = 10:20
y > 15
y %% 2 == 0 # is TRUE for those elements of y that are an even number
#> task
```


b) Return a logical vector that returns a TRUE for all elements of y that are <= 14.

```{r "6 b)"}
y <= 14
```

c) Return a logical vector that returns a TRUE for all elements of y that are smaller equal 14 or bigger equal 18

```{r "6 c)"}
y <= 14 | y >= 18
#< hint
display("Use the logical OR | to combine conditions of the form y <= ??? and y >= ???")
#>
```

d) Return a logical vector that returns a TRUE for all elements of y that are between 15 and 19 (including 15 and 19)

```{r "6 d)"}
y >= 15 & y <= 19
#< hint
display("Use the logical AND & to combine conditions of the form y >= ??? and y <= ???")
#>
```

e) We can use logical vectors to index vectors. For example, to get all values of y that are strictly above 15, we can type

```{r "6 e)"}
#< task
y[y>15]
#> task
```

f) Show all values of y that are strictly below 13

```{r "6 f)"}
y[y<13]
```

g) Show all values of y that are between 14 and 18

```{r "6 g)"}
y[y>=14 & y <= 18]
```

h) Note that we can also store a logical vector in a variable

```{r "6 h)"}
#< task notest
ind = (y >=18)
ind
y[ind]
```

the function which gets the index of all elements in which a logical vector is TRUE

```{r "6 h) 2"}
which(ind)
#> task
#<
give.award("It's only logical","You learned about the logical operators & (and) and | (or) and how you can use them to select rows from a vector.")
#>
```

## Exercise 7 -- Lists

Besides vectors we often encounter in R the following objects: list, data.frame and matrix. Let us briefly get used to lists here.

a) A list is a very flexible container that can contain variables of different kinds, like numbers, characters, vectors or lists itself

```{r "7 a)"}
#< task
x = list(a="Hi",b=1:3,c=list(y=3,z=4))
x
```

The elements of list often have a name. We can access an element of the list by its name or index in several ways:

```{r "7 a) 2"}
x[[1]]
x[["a"]]
x$a
#> task
```

b) Show the element b of x using the $ operator

```{r "7 b)"}
x$b
```

c) Show the element c of x using its index, i.e. 3, and the [[...]] notation

```{r "7 c)"}
x[[3]]
```

d) We can also simply add or change elements of an list

```{r "7 d)"}
#< task
x$d = "New element" 
x
#> task
```

e) Add to x the element "e" with the value 100

```{r "7 e)"}
x$e = 100
```

f) Note: if we index elements of a list with the [...] notation, we get a sublist which is itself a list.

```{r "7 f)"}
#< task notest
x[[1]] # yields the element a of x
x[1] # yields a list with one element $a
x[1:2] # yields a list with elements $a and $b
#> task
#<
give.award("list-ener","You heard about lists...")
#>
```

## Exercise 8 -- Data frames

a) A data frame is a list whose elements are vectors of the same length. A data frame is the common way to store data in R. 

```{r "8 a)"}
#< task
df = data.frame(a=5:7, b=4, c=c("A","B","C"), d = runif(3))
df
#> task
```

Note that b is automatically changed to a column of 3 times 4.

b) If we use the $ notation on a data.frame, we get the corresponding column. Show the column 'a' of df 

```{r "8 b)"}
df$a
```

c) If we want to get the element in the 2nd row in column a of df, we can have 4 different ways:

```{r "8 c)"}
#< task notest
# Subset data.frame like lists
df$a[2]
df[["a"]][2]

# Subset data.frame like a matrix
df[2,"a"]
df[2,1]
#> task
```

d) Store in the variable x the rows 2 and 3 of column d in df and then show x

```{r "8 d)"}
x = df$d[2:3]
x
```

We will learn much more about data frames in later problem sets.

```{r "8 d) 2"}
#<
give.award("framing the data","R users simply love to put their data into frames.")
#>
```

## Exercise 9 -- Matrices

a) A matrix works for a user similar like a data frame, but all its columns must be the same type (usually numbers). Internally, a matrix is stored differently than a data frame and some sort of computations can be performed much quicker.

```{r "9 a)"}
# Generate a matrix with 3 rows and 2 columns
#< task
m = matrix(1:6,nrow=3,ncol=2)
#< task notest
m
```

b) There are a lot of functions for matrices, e.g. taking the transpose

```{r "9 b)"}
t(m)
```

c) Similar as with vectors we can do a lot of element wise computations on matrices

```{r "9 c)"}
m * 3
#> task
```

d) Compute the sum of the matrix m and m

```{r "9 d)"}
m+m
```

e)

```{r "9 e)"}
#< task notest
m*m # elementwise multiplication
t(m) %*% m # matrix multiplication
#> task
```


We will learn more about computing with matrices in later problem sets.

```{r "9 e) 2"}
#<
give.award("Intro the matrix","Tired of only working with vectors? A matrix opens up a whole new dimension for you!")
#>
```


## Ignore

```{r "9 e) 3"}
library(RTutor);library(restorepoint)
#setwd("D:/lehre/empIOUlm/rtutor")
ps.name = "r_intro"; sol.file = paste0(ps.name,"_sol.Rmd")
name.rmd.chunks(sol.file, only.empty.chunks=FALSE)
create.ps(sol.file=sol.file, ps.name=ps.name,libs=NULL)

```
