# Problemset DataFramesExample


```{r 'check_ps', include=FALSE}

user.name = '' # set to your user name
# To check your problem set, save your file (Ctrl-S) and then run the RStudio Addin 'Check Problemset'

# Alternatively run the following lines 
library(RTutor)
ps.dir = getwd(); # directory of this file
ps.file = 'DataFramesExample.Rmd' # name of this file
check.problem.set('DataFramesExample', ps.dir, ps.file, user.name=user.name, reset=FALSE)
```


Example of data frame printing in an RTutor Problemset

#### Author:  Sebastian Kranz

Adapt the working directory below and press Ctrl-Alt-R (run all chunks). This creates the problem set files and the sample solution from this solution file.



## Exercise 1 -- Table outputs

The shiny based inteface of RTutor shows data.frames by default as html tables. This example illustrates how, you can adapt the settings of how data.frames are shown.

Use default settings for printing data.frames, that may have been adapted in the call to `show.ps`.
```{r "1",optional=TRUE}
T = 40
data.frame(x=runif(T),y=runif(T)*100000)
2*2
```

You can overwrite the default settings for specific chunks, by specifying chunk options.
```{r "1__2",signif.digits=5, round.digits=3, table.max.rows=30, html.data.frame=FALSE, optional=TRUE}
T = 40
data.frame(x=runif(T),y=runif(T)*100000, z=runif(T))
```

---

## Exercise 2 -- Empty


Enter 2*2
```{r "2___5143129"}

```

That's all folks!
