# Problemset DataFramesExample

Example of data frame printing in an RTutor Problemset

#### Author:  Sebastian Kranz

Adapt the working directory below and press Ctrl-Alt-R (run all chunks). This creates the problem set files and the sample solution from this solution file.

#< ignore
```{r "create_ps",include=FALSE , eval=!TRUE}
library(RTutor)
#library(restorepoint)
#setwd("D:/libraries/RTutor/examples")
setwd("D:/libraries/RTutor/RTutor/inst/examples")
ps.name = "DataFramesExample"; sol.file = paste0(ps.name,"_sol.Rmd")
libs = c() 
# Important add the parameter addons="quiz" !
create.ps(sol.file=sol.file, ps.name=ps.name, user.name=NULL,libs=libs, stop.when.finished=FALSE)

# You can set default parameters for showing data frames here
show.ps(ps.name, user.name="John Does", round.digits=4, table.max.rows=20)

stop()
```

Load libraries so that knit HTML also works
```{r}
library(yaml)
library(magrittr)
library(leaflet)
```

#>


## Exercise 1 -- Table outputs

The shiny based inteface of RTutor shows data.frames by default as html tables. This example illustrates how, you can adapt the settings of how data.frames are shown.

Use default settings for printing data.frames, that may have been adapted in the call to `show.ps`.
```{r "mychunk", optional=TRUE}
#< task_notest
T = 40
data.frame(x=runif(T),y=runif(T)*100000)
2*2
#>
```

You can overwrite the default settings for specific chunks, by specifying chunk options.
```{r "mychunk2", signif.digits=5, round.digits=3, table.max.rows=30, html.data.frame=FALSE, optional=TRUE}
#< task_notest
T = 40
data.frame(x=runif(T),y=runif(T)*100000, z=runif(T))
#>
```

---

## Exercise 2 -- Empty


Enter 2*2
```{r}
2*2
```

That's all folks!
