# Problemset Example


```{r 'check_ps', include=FALSE}

user.name = 'Jane Doe' # set to your user name
# To check your problem set, save your file (Ctrl-S) and then run the RStudio Addin 'Check Problemset'

# Alternatively run the following lines 
library(RTutor)
ps.dir = getwd(); # directory of this file
ps.file = 'Example.Rmd' # name of this file
check.problem.set('Example', ps.dir, ps.file, user.name=user.name, reset=FALSE)
```


Example of an RTutor interactive Problemset
Author:  Sebastian Kranz
Date:    12.06.2014

Adapt the working directory below and press Ctrl-Alt-R (run all chunks). This creates the problem set files and the sample solution from this solution file.

## Exercise 1 -- Summary statistics

a) First part

Some Math:

$$x = \beta^2$$

info("Some math in an info block") # Run this line (Strg-Enter) to show info


We often want to compute some summary statistic of a vector.
For example:


```{r "1_a",fig.height=3}
x = 10:20
# Computing the sum of x
sum(x)
# Generating a data frame
df = data.frame(a=c(1,1,2,2,2,3,3,4), b = runif(8), c="Hi")
df
```

Now compute the mean of x.
```{r "1_a_2"}
mean(x)
```


info("useful functions for numeric vectors") # Run this line (Strg-Enter) to show info



## Exercise 2 -- Computing with vectors
a) Let y be a vector that contains the squared elements of x. Show y
```{r "2_a"}
y = x^2
y
```
