
---
title: Problem Set Example
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE, echo=FALSE}
# Load libraries and source extra code
library(leaflet)
library(RTutor)


# options for rendering data frames
RTutor::set.knit.print.opts(html.data.frame=FALSE, table.max.rows=25, round.digits=8, signif.digits=8)

# continue knitting even if there is an error
knitr::opts_chunk$set(error = TRUE) 
```

# Problemset Example

Example of an RTutor interactive Problemset
Author:  Sebastian Kranz
Date:    12.06.2014

Adapt the working directory below and press Ctrl-Alt-R (run all chunks). This creates the problem set files and the sample solution from this solution file.

## Exercise 1 -- Summary statistics

a) First part

Some Math:

$$x = \beta^2$$


***

### Info: Some math in an info block
$$y = \alpha^2$$

***



We often want to compute some summary statistic of a vector.
For example:


```{r "1_a",fig.height=3}
x = 10:20
# Computing the sum of x
sum(x)
# Generating a data frame
df = data.frame(a=c(1,1,2,2,2,3,3,4), b = runif(8), c="Hi")
df
```

Now compute the mean of x.
```{r "1_a_2"}
mean(x)
```



***

### Info: useful functions for numeric vectors
Here are examples for useful R functions
```{r "1_a_3"}
max(c(1,5,2)) # returns maximum
min(c(1,5,2)) # returns minimum

sum(c(1,5,2,NA), na.rm=TRUE) # returns sum of all numbers, ignore NA
cumsum(c(1,5,2)) # returns cummulated sum
diff(c(1,5,2)) # returns the vector of differences
```

***




***

### Award: mean means mean
Well, in some occasions one can just guess the name of an R function. The function to compute the mean of a vector, or matrix is called 'mean'. Usually, it is much quicker to goggle than to guess function names, however.

***


## Exercise 2 -- Computing with vectors
a) Let y be a vector that contains the squared elements of x. Show y
```{r "2_a_4"}
y = x^2
y
```
