# Problemset HtmlWidget Example

An example how to show an HTML widget from the leaflet package.

Adapt the working directory below and press Ctrl-Alt-R (run all chunks). This creates the problem set files and the sample solution from this solution file.

#< ignore
```{r "create_ps",include=FALSE , eval=!TRUE}
library(RTutor)
setwd("D:/libraries/RTutor/RTutor/inst/examples")
ps.name = "HtmlWidgetExample"; sol.file = paste0(ps.name,"_sol.Rmd")
libs = c("leaflet","magrittr","network3d") 
# Important add the parameter addons="quiz" !
create.ps(sol.file=sol.file, ps.name=ps.name, libs=libs, addons="quiz")
show.ps(ps.name)

stop()
```

Load libraries so that knit HTML also works
```{r "load_libs"}
library(yaml)
library(magrittr)
library(leaflet)
```

#>


## Exercise 1 -- Leaflet

Check to see an interactive leaflet centered on the department of Mathematics and Economics at Ulm University. The place where RTutor has been created.

To show an html widget output, you must add the chunk option `output="htmlwidget"`. Further you must set the option `widget` to the package of the HtmlWidget you want to show, here it is `leaflet`.
```{r output="htmlwidget", widget="leaflet"}
#< task_notest
SFmap <- leaflet() %>% 
  addTiles() %>% 
  setView(9.9620289,48.4251394, zoom = 11) %>% 
  addMarkers(9.963614702355699,48.42517242865929, popup = 'RTutor was created here at Ulm University')
SFmap
#>
```

Here you can see what else you can do with the great leaflet packages:

http://rstudio.github.io/leaflet/

Here is a list of some other htmlwidgets, which now can also be used with RTutor:

http://www.htmlwidgets.org/
