# Problemset Example

Example of an RTutor interactive Problemset

#### Author:  Sebastian Kranz

Adapt the working directory below and press Ctrl-Alt-R (run all chunks). This creates the problem set files and the sample solution from this solution file.

#< ignore
```{r "create_ps",include=FALSE , eval=!TRUE}
library(RTutor)
#library(restorepoint)
#setwd("D:/libraries/RTutor/examples")
setwd("D:/libraries/RTutor/RTutor/inst/examples")
ps.name = "LeafletExample"; sol.file = paste0(ps.name,"_sol.Rmd")
libs = c("leaflet","magrittr") 
# Important add the parameter addons="quiz" !
create.ps(sol.file=sol.file, ps.name=ps.name, user.name=NULL,libs=libs, stop.when.finished=FALSE, addons="quiz")

show.ps(ps.name, user.name="John Does")

stop()
```

Load libraries so that knit HTML also works
```{r}
library(yaml)
library(magrittr)
library(leaflet)
```

#>


## Exercise 1 --Leaflet


Check to see an interactive leaflet centered on the department of Mathematics and Economics at Ulm University. The place where RTutor has been created.
```{r results='asis', output="htmlwidget", widget="leaflet"}
#< task_notest
SFmap <- leaflet() %>% 
  addTiles() %>% 
  setView(9.9620289,48.4251394, zoom = 11) %>% 
  addMarkers(9.963614702355699,48.42517242865929, popup = 'RTutor was created here at Ulm University')
SFmap
#>
```

Here you can see what else you can do with the great leaflet packages:

http://rstudio.github.io/leaflet/

Here is a list of some other htmlwidgets, which now can also be used with RTutor:

http://www.htmlwidgets.org/

---

## Exercise 2 -- Empty


Enter 2*2
```{r}
2*2
```

That's all folks!
