# Problemset Example

Example of an RTutor interactive Problemset

#### Author:  Sebastian Kranz

Adapt the working directory below and press Ctrl-Alt-R (run all chunks). This creates the problem set files and the sample solution from this solution file.

#< ignore
```{r "create_ps",include=TRUE , eval=TRUE}
library(RTutor)
library(yaml)
#library(restorepoint)
#setwd("D:/libraries/RTutor/examples")
setwd("D:/libraries/RTutor/RTutor/inst/examples")
ps.name = "QuizExample"; sol.file = paste0(ps.name,"_sol.Rmd")
libs = c("yaml") 
# Important add the parameter addons="quiz" !
create.ps(sol.file=sol.file, ps.name=ps.name, user.name=NULL,libs=libs, stop.when.finished=FALSE, addons="quiz")

show.ps(ps.name, user.name="John Does")

stop()
```
#>

## Exercise 1 -- Quiz

### A single choice quiz

#< quiz "single"
question: What is 20*20?
sc:
    - 100
    - 200
    - 400*
    - 500
success: Great, your answer is correct!
failure: Try again.
#>

### A multiple choice quiz

#< quiz "multi"
question: Check all prime numbers
mc:
    - 2*
    - 3*
    - 4
    - 5*
    - 6
    - 7*
    - 8
success: Great, all answers are correct!
failure: Not all answers correct. Try again.
#>


### An open question

#< quiz "music"
question: Johann ???? Bach
answer: Sebastian
#>

### An open question with a numeric answer

#< quiz "pi"
question: 2. State pi up to 2 digits
answer: 3.14
roundto: 0.01
#>


### A quiz consisting of several questions (and a single check button)

#< quiz "parts"
parts:
  - question: 1. What is 20*20?
    choices:
        - 100
        - 200
        - 400*
        - 500
    multiple: FALSE
    success: Great, your answer is correct!
    failure: Try again.
  - question: 2. State pi up to 2 digits
    answer: 3.14
    roundto: 0.01
#>


#< award "Quiz master"
You solved the quiz!
#>


Enter 2*2
```{r}
2*2
```


---

## Exercise 2 -- Empty


Enter 2*2
```{r}
2*2
```

That's all folks!
